\chapter{Introduction to Arduino programming}

\section{Sketches}

The Arduino programs are commonly referred to as \textit{sketches}. A sketch is not a single file, but an entire directory, usually populated by files with the \texttt{.ino} extension, contaning the source code of the program. \texttt{.cpp}, \texttt{.c}, \texttt{.S} (assembly files), and header files like \texttt{.h} are also accepted as files contaning source code. Every sketch must contain a \texttt{.ino} file with a file name matching the sketch root directory name \cite{arduino_sketch_specs}.

The Arduino programming language, in which \texttt{.ino} files are written, is a C++ dialect with very subtle differences, mostly concerning linking of files and ommitting function declarations. Bodies of function definitions and variable declarations can be considered as a perfectly valid C/C++ code. 

The essential functions that an Arduino programmer has to define in their sketch are \texttt{void setup()} and \texttt{void loop()}. The \texttt{setup()} function is called once, when the program is booted. That is when the Arduino board is powered up, the program is uploaded to it, or when the reset button is pressed. The \texttt{loop()} function is called in a loop indefinitely. This is the proper place to define any interations with the board, detecting inputs, updating timers, etc.

\section{Arduino board interface}

The target board \texttt{Arduino Uno R3 SMD} provides 6 analog and 14 digital general purpose pins, some of which can be used for I2C and SPI protocol communication, or PWM modulation. The most simple to use and simulate are the digital pins for general purpose I/O, so this thesis focuses exlucively on using these pins as such \cite{arduino_r3_smd_specs}.

The digital pins can be manipulated using only these 3 functions:
\begin{itemize}
    \item \texttt{void pinMode(pin, mode)} - configures specified pin either as an input or an output
    \item \texttt{value digitalRead(pin)} - reads the voltage on the pin, stores it as a digitla value, either 1 or 0
    \item \texttt{void digitalWrite(pin, value)} - when set to output, sets the voltage of the pin to corresponding value. If set to input, current driven components may not work correctly, since the current is suppressed by an internal pull-up resistor. 
\end{itemize}

Pins are identified by their index. \texttt{mode} and \texttt{value}'s underlying type is also an integer, but it is preferred to use predefined constants, such as \texttt{INPUT}/\texttt{OUTPUT}, \texttt{HIGH}/\texttt{LOW} and such. Below is an example of a sketch from the Arduino's documentation, using only the described functions:

\begin{lstlisting}[language=C, caption=Sketch example \cite{arduino_digital_read}]
int ledPin = 13;  // LED connected to digital pin 13
int inPin = 7;    // pushbutton connected to digital pin 7
int val = LOW;    // variable to store the read value

void setup() {
  pinMode(ledPin, OUTPUT);  // sets the digital pin 13 as output
  pinMode(inPin, INPUT);    // sets the digital pin 7 as input
}

void loop() {
  val = digitalRead(inPin);   // read the input pin
  digitalWrite(ledPin, val);  // sets LED to the button's value
}
\end{lstlisting}

\section{Build process}

Before the code is compiled, all \texttt{.ino} go through a simple preprocessing stage. \texttt{\#line} directives are added. These link to the original line and file, when the code is debugged. All files are concatenated together, function declarations are generated at the top, and the \texttt{Arduino.h} is included. This header file is crucial as it declares the \texttt{void} and \texttt{loop} functions, functions for manipulating pins, common constants and mathematical functions, etc. At this point the sketch becomes a complete C++ source code.

It is also at this where we steer away from the standard compilation process, and apply our own \texttt{Arduino.h} implementation in our simulator. Usually, the C++ code would be linked to an implementation that contains binaries and constants for the given microchip installed on the target board. Then it would be compiled with a compiler for the given instruction set (e.g. \texttt{avr-gcc}). The code would be then transmitted over USB to the user's Arduino.

In our case, we include the entire preprocessed sketch in our simulator project written in C++, and compile it as whole for the instruction set of the host machine. This executable will implement mock functions declared in the \texttt{Arduino.h}, arduino's internal state, and a way to communicate the change of state to an external GUI application.

The entire process from preprocessing to compilation, can be orchestrated with the use of \texttt{arduino-cli} command line application. We will use it to preprocess the \texttt{.ino} for us, so that we end up with a clean C++ file we can include in our project. 

\section{Specific goals of the thesis}

\subsubsection{Target platforms:}
\begin{itemize}
    \item Linux (x86-64)
    \item Windows 10 or higher (64-bit)
\end{itemize}

\subsubsection{Software prerequisits:}
\begin{itemize}
    \item Minimal system requirements of Avalonia UI
    \item arduino-cli
    \item .NET 9 runtime or higher
    \item C++ compiler
    \item C++ Boost libraries:
    \begin{itemize}
        \item asio
        \item interprocess
        \item program\_options
    \end{itemize}
\end{itemize}

\subsubsection{Simulator functional requirements}
\begin{itemize}
    \item Defines platform independent constants and macros from \texttt{Arduino.h}.
    \item Implements \texttt{digitalWrite}, \texttt{digitalRead}, \texttt{pinMode} and \texttt{millis} functions from \texttt{Arduino.h}.
    \item Declares and calls \texttt{setup} and \texttt{loop} functions with the behavior expected on an Arduino machine.
    \item Compiles and runs any valid \texttt{Arduino UNO R3 SMD} sketch, preprocessed with the \texttt{arduino-cli} tool, containing only symbols mentioned above.
    \item Reflects behavior of the above functions in an internal Arduino state. Logs changes of this state in a text file.
    \item Integrates an interprocess communication mechanism for exchanging data with an external GUI application.
    \item The events incoming from a GUI application must be resolved in a manner, that preserves the continuity of events.
\end{itemize}

\subsubsection{GUI functional requirements}
\begin{itemize}
    \item The application must stay responsive during debugging of the simulator process.
    \item The application must be able to connect to the interprocess mechanism of the simulator.
    \item Simulates simplified digital circuits. Exact physical properties, like voltages and currents might not be implemented. Built-in components may be implicitly connected to ground.
    \item Implements mechanism for defining custom components and visuals with \texttt{.yaml}, \texttt{.cs} and \texttt{.svg} files. These components include interactive controls, such as buttons.
    \item Implements mechanism for defining a circuit (scene) using the custom components.
    \item Predefines LED, button, voltage source and Arduino components, and a scene which utilizes these components.
    \item The changes to the simulator's internal state must be immediately reflected on visual components connected to the Arduino in the scene.
    \item The scene is able to be zoomed and panned. The visuals of the components must be rendered using vector graphics.
    \item Multiple interactable components may be pressed at once, with behavior expected from the source code of the sketch.
\end{itemize}